---
title: "Validation Accuracy Analysis on imdb movie review data"
output:
  word_document: default
  html_notebook: default
---

```{r, results='hide'}
library(keras)

imdb <- dataset_imdb(num_words = 10000)
c(c(train_data, train_labels), c(test_data, test_labels)) %<-% imdb

#Padding list of integers to tensors and dividing the training set to training and validation sets

vectorize_sequences <- function(sequences, dimension = 10000) {
  # Create an all-zero matrix of shape (len(sequences), dimension)
  results <- matrix(0, nrow = length(sequences), ncol = dimension)
  for (i in 1:length(sequences))
    # Sets specific indices of results[i] to 1s
    results[i, sequences[[i]]] <- 1
  results
}

# Our vectorized training data
x_train <- vectorize_sequences(train_data)
# Our vectorized test data
x_test <- vectorize_sequences(test_data)

#We should also vectorize our labels, which is straightforward:
# Our vectorized labels

y_train <- as.numeric(train_labels)
y_test <- as.numeric(test_labels)

#In order to monitor during training the accuracy of the model on data that it has never seen before, we will create a "validation set" by setting apart 10,000 samples from the original training data:

val_indices <- 1:10000

x_val <- x_train[val_indices,]
partial_x_train <- x_train[-val_indices,]

y_val <- y_train[val_indices]
partial_y_train <- y_train[-val_indices]

```

#Implementing text classification using keras with simple network
```{r}
library(keras)

model <- keras_model_sequential() %>% 
  layer_dense(units = 16, activation = "relu", input_shape = c(10000)) %>% 
  layer_dense(units = 16, activation = "relu") %>% 
  layer_dense(units = 1, activation = "sigmoid")

model %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("accuracy")
)

history <- model %>% fit(
  partial_x_train,
  partial_y_train,
  epochs = 20,
  batch_size = 512,
  validation_data = list(x_val, y_val)
)

History<-data.frame(history$metrics)
plot(history)
```


For the IMDB example that we discussed in class, do the following:

	1. Try using one or three hidden layers, and see how doing so affects validation and test accuracy.
	
```{r}
library(keras)

model_Layers <- keras_model_sequential() %>% 
  layer_dense(units = 16, activation = "relu", input_shape = c(10000)) %>%  
  layer_dense(units = 16, activation = "relu") %>% 
  layer_dense(units = 16, activation = "relu") %>%
  layer_dense(units = 16, activation = "relu") %>%
  layer_dense(units = 1, activation = "sigmoid")

model_Layers %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("accuracy")
)

history_Layers <- model_Layers %>% fit(
  partial_x_train,
  partial_y_train,
  epochs = 20,
  batch_size = 512,
  validation_data = list(x_val, y_val)
)
Hist_Layers<-data.frame(history_Layers$metrics)
plot(history_Layers)
```


	2. Try using layers with more hidden units or fewer hidden units: 32 units, 64 units, and so on.
```{r}
library(keras)

model_32Units <- keras_model_sequential() %>% 
  layer_dense(units = 16, activation = "relu", input_shape = c(10000)) %>% 
  layer_dense(units = 32, activation = "relu") %>% 
  layer_dense(units = 1, activation = "sigmoid")

model_32Units %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("accuracy")
)

history_32Units <- model_32Units %>% fit(
  partial_x_train,
  partial_y_train,
  epochs = 20,
  batch_size = 512 ,
  validation_data = list(x_val, y_val)
)
Hist_32Units<-data.frame(history_32Units$metrics)
plot(history_32Units)
```


```{r}

library(keras)

model_64Units <- keras_model_sequential() %>% 
  layer_dense(units = 16, activation = "relu", input_shape = c(10000)) %>% 
  layer_dense(units = 64, activation = "relu") %>%
  layer_dense(units = 64, activation = "relu") %>%
  layer_dense(units = 1, activation = "sigmoid")

model_64Units %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("accuracy")
)

history_64Units <- model_64Units %>% fit(
  partial_x_train,
  partial_y_train,
  epochs = 20,
  batch_size = 512 ,
  validation_data = list(x_val, y_val)
)
Hist_64Units<-data.frame(history_64Units$metrics)
plot(history_64Units)
```


	3. Try using the mse loss function instead of binary_crossentropy. 
```{r}
library(keras)

model_LF <- keras_model_sequential() %>% 
  layer_dense(units = 16, activation = "relu", input_shape = c(10000)) %>% 
  layer_dense(units = 16, activation = "relu") %>% 
  layer_dense(units = 1, activation = "sigmoid")

model_LF %>% compile(
  optimizer = "rmsprop",
  loss = "mse",
  metrics = c("accuracy")
)

history_LF <- model_LF %>% fit(
  partial_x_train,
  partial_y_train,
  epochs = 20,
  batch_size = 512,
  validation_data = list(x_val, y_val)
)
Hist_LF<-data.frame(history_LF$metrics)
plot(history_LF)
```
	
	4. Try using the tanh activation (an activation that was popular in the early days of neural networks) instead of relu.
```{r}
library(keras)

model_AF <- keras_model_sequential() %>% 
  layer_dense(units = 16, activation = "tanh", input_shape = c(10000)) %>% 
  layer_dense(units = 16, activation = "tanh") %>% 
  layer_dense(units = 1, activation = "sigmoid")

model_AF %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("accuracy")
)

history_AF <- model_AF %>% fit(
  partial_x_train,
  partial_y_train,
  epochs = 20,
  batch_size = 512,
  validation_data = list(x_val, y_val)
)
Hist_AF<-data.frame(history_AF$metrics)
plot(history_AF)
```
	
	5. Use any technique we studied in class, and these include regularization, dropout, etc., to get your model to perform better on validation.
	
	##With Dropout Layer
```{r}
model_Drop <- keras_model_sequential() %>% 
  layer_dense(units = 16, activation = "relu", input_shape= c(10000)) %>% layer_dropout(rate = 0.15) %>%
  layer_dense(units = 16, activation = "relu") %>% layer_dropout(rate = 0.15) %>% 
  layer_dense(units = 1, activation = "sigmoid")

model_Drop %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("accuracy")
)

history_Drop <- model_Drop %>% fit(
  partial_x_train,
  partial_y_train,
  epochs = 20,
  batch_size = 512,
  validation_data = list(x_val, y_val)
)
Hist_Drop<-data.frame(history_Drop$metrics)
plot(history_Drop)
```
	


	##With Ridge Regression as penalty
```{r}
model_Ridge <- keras_model_sequential() %>% 
  layer_dense(units = 16,kernel_regularizer = regularizer_l2(0.00005),activation = "relu", input_shape= c(10000)) %>% 
  layer_dense(units = 16,kernel_regularizer = regularizer_l2(0.00005), activation = "relu") %>%
  #layer_dense(units = 8,kernel_regularizer = regularizer_l2(0.001), activation = "relu") %>%
  layer_dense(units = 1, activation = "sigmoid")

model_Ridge %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("accuracy")
)

history_Ridge <- model_Ridge %>% fit(
  partial_x_train,
  partial_y_train,
  epochs = 20,
  batch_size = 512,
  validation_data = list(x_val, y_val)
)
Hist_Ridge<-as.data.frame(history_Ridge$metrics)
plot(history_Ridge)
```



	##With Lasso as penalty
```{r}
model_Lasso <- keras_model_sequential() %>% 
  layer_dense(units = 16,kernel_regularizer = regularizer_l1(0.0001),activation = "relu", input_shape= c(10000)) %>% 
  layer_dense(units = 16,kernel_regularizer = regularizer_l1(0.0001), activation = "relu") %>%
  #layer_dense(units = 8,kernel_regularizer = regularizer_l1(0.0001), activation = "relu") %>%
  layer_dense(units = 1, activation = "sigmoid")

model_Lasso %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("accuracy")
)

history_Lasso <- model_Lasso %>% fit(
  partial_x_train,
  partial_y_train,
  epochs = 20,
  batch_size = 512,
  validation_data = list(x_val, y_val)
)
Hist_Lasso<-data.frame(history_Lasso$metrics)
plot(history_Lasso)
```

#Data_Visualization	
```{r}

Models_Val_Accuracy<-data.frame(epochs=1:10,History[1:10,4],Hist_32Units[1:10,4],Hist_64Units[1:10,4],Hist_Layers[1:10,4],Hist_AF[1:10,4],Hist_LF[1:10,4],Hist_Drop[1:10,4],Hist_Ridge[1:10,4],Hist_Lasso[1:10,4])

colnames(Models_Val_Accuracy)<-c("epochs","Model Val Acc","Units 32","Units 64","Hidden Layers 3","Tanh AF","MSE LF","DrouOut","Ridge regularizer","Lasso regularizer")

library(data.table)
Model_ValAcc_transpose <- transpose(Models_Val_Accuracy)
rownames(Model_ValAcc_transpose)<-colnames(Models_Val_Accuracy)
colnames(Model_ValAcc_transpose)<-rownames(Models_Val_Accuracy)

library(dplyr)
Max_Model_Val_Accuracy<-summarise_each(Models_Val_Accuracy, funs(max))
```

